---
title: "hw02-MeiqiYu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gapminder)
library(tidyverse)
```

# Smell test the data
* Gapminder is a data.frame which is a rectangular collection of variables (in the columns) and observations (in the rows). 
* It contains 6 cloumns and 1704 rows.
* We could get the facts about "extent" and "size" in the following six ways.
1. `gapminder`
It shows the entire dataframe which contains 6 comlumns and 1704 rows. 
```{r}
gapminder
```

2. `head(gapminder)` 
By runnig this command, it is clear that this datframe contains 6 rows as it shows below.
```{r}
head(gapminder)
```

3. `str(gapminder)` 
The result shows apparently the class, number of rows and varibles.
```{r}
str(gapminder)
```

4. `summary(gapminder)` 
It is a summary about the dataframe so you can easily find that how many varibles it contains. It will be used to help us get a knowledge about the minimun value, the max value, 1st Qu.,3rd Qu. and median.
```{r}
summary(gapminder)
```
5.`dim(gapminder)` 
`dim()` can be used to obtain the dimensions of the data frame (number of rows and number of columns). The output is a vector.
```{r}
dim(gapminder)
```
6.`nrow()`and`ncol()` 
These two function can help us get number of rows and number of columns, respectively.
```{r}
nrow(gapminder)
ncol(gapminder)
```
 

* There are six varibles in gapminder: country, continent, year, lifeExp, pop and gdpRercap. 
Data type of each varible:
 * country: character
 * continent: character
 * year: integar
 * lifeExp: double
 * pop: integar
 * gdpPercap: double
 The first two varibles are categerical varibles(qualitative) and the last four are quantitative varibles(numerical).
 
# Explore individual variables
In this part, I choose the following two varibles to explore: `continent` and `year`.
* Possible range of `continent`: 
Asia, Europe, Africa   Americas and Oceania 
* Possible values of `year`:
1952 1957 1962 1967 1972 1977 1982 1987 1992 1997 2002 2007.
```{r}
unique(gapminder$continent) 
unique(gapminder$year) 
```

* Typical values contain the Min., 1st Qu., Median, Mean, 3rd QU., Max. For the `year` varible, the Min= 1952, 1st quartile= 1966, Median= 1980, Mean= 1980, 3rd quartile= 1993, Max= 2007.
```{r}
summary(gapminder$year)
```
* Spread of data 

Overall range = max-min
Overall range of `year` equals to 55.

* Distribution of data 

 * The shape of `year` is uniform, which means about the same amount of data occurs for each varible value.
 * Mean and Median of `year` both equal to 1980.
```{r}
ggplot(gapminder,aes(year))+
  geom_histogram(bins=50)
```
 
 * The standard deviation of an observation variable is the square root of its variance. We can know the standard deviation of `year` equals to 17.26533 from the below.
```{r}
sd(gapminder$year)
```

# Explore various plot types

* A scatterplot of lifeExp and gdpPercap in blue.
```{r}
ggplot(gapminder,aes(x=lifeExp, y=(gdpPercap)))+
  geom_point(color='steelblue',size=1)
 
```


* A histogram of lifeExp with bins = 100.

```{r}
ggplot(gapminder,aes(lifeExp))+
  geom_histogram(bins=100)
```

* A densityplot of lifeExp.
```{r}
ggplot(gapminder,aes(lifeExp))+
  geom_density(bw=2)
```

*A boxplot for all the continents vs lifeExp.


```{r}
a <- ggplot(gapminder,aes(continent,lifeExp)) + 
  scale_y_log10()
a + geom_boxplot()
```



*A violin plot for all entries of Americas and Africa occuring in the '60s.   vs lifeExp.

```{r}
gapminder %>% 
  filter((continent == 'Americas' |continent == 'Africa'|continent == 'Asia') & year>=1960 & year < 1970) %>% 
  ggplot(aes(continent,lifeExp))+
  geom_violin()
```

# Use filter(), select() and %>%

```{r}
select(gapminder, year,lifeExp) %>% 
  filter(year>=1960 & year < 1970) %>% 
  ggplot(aes(year,lifeExp))+
  geom_line()
```

# But I want to do more!
By running this code `filter(gapminder, country == c("Rwanda", "Afghanistan"))`, the analyst didn't secceed to get what they want. They only got part of the result. 
```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```
The correct way to do this is listed below:
```{r}
  d <- filter(gapminder,country == 'Rwanda' | country == 'Afghanistan')
  knitr::kable(d)
```






